Thermodynamic response functions in a cell fluid model

### 2024-09-15 paper_text.tex
make references more consistent
replace dependency with dependence; fix references order

\subsection{Pressure as a function of temperature and density}
In this work, by density we mean the particle number density $\rho = \langle N \rangle / V$.
In the framework of the grand canonical ensemble the average number of particles $\left\langle N \right\rangle$ is found by 
\begin{eqnarray*}
	\langle N \rangle & = & \left(\frac{\partial \ln\Xi}{\partial \beta \mu}\right)_{T,V} 
	\\
	& = & V \left(\frac{\partial P}{\partial \mu}\right)_{T,V}.
\end{eqnarray*}
From this equation it follows
\begin{equation*}
	\frac{\langle N \rangle}{V} = \frac{\langle N \rangle}{v N_v} = \left(\frac{\partial P}{\partial \mu}\right)_{T,V}
\end{equation*}
and thus
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = v \left(\frac{\partial P}{\partial \mu}\right)_{T,V} 
	= \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
The quantity $\rho^*$, on the one hand, is the reduced particle number density, which is the notation commonly used in the literature on simple fluids~\cite{HansenMcDonald2013}. In the context of the cell model, on the other hand, it is the average number of particles per cell.\footnote{In our previous works we denoted the reduced number density by $\bar{n}$. In the current work we switch to more common notation $\rho^*$.}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = & g_1 - \frac{g_3}{g_4}\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right) - \frac{1}{6} \frac{g_3^3}{g_4^2} 
	\nonumber\\
	& = & g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is
\begin{equation*}
	\rho^*_c = 0.978.
\end{equation*}
Equations~\eqref{eq:density} and~\eqref{eq:eosMT} jointly define a parametric relationship between pressure and density, with $M$ serving as the parameter. Figure~\ref{fig1} shows the dependency of the reduced pressure $P^*$ on the density $\rho^*$ for various values of the reduced temperature $\tau$, at and above the critical temperature.

This dependency can also be expressed explicitly. To achieve this, we combine the equations~\eqref{eq:density} and~\eqref{eq:ro_M} to express the effective chemical potential $M$ as a function of density and temperature:
\begin{equation}\label{eq:M_nT}
	\bar M = \frac{\rho_{n}}{\beta W(0)} - (\rho^* - \rho^*_c),
\end{equation}
where
\begin{align} \label{eq:ro_nT}
	& \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
	& \alpha_n = \arccos \left[ \left( - \frac{9 g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (\rho^*_c - \rho^*)\right]. \nonumber 
\end{align}
The notation $\bar{M}$ represents the effective chemical potential as a function of temperature and density $\rho^*$, while $M$ represents the effective chemical potential as a function of temperature and chemical potential~$\mu$.

@@

Thus, in this Section, we presented two forms of the equation of state. The first one expresses the pressure as a function of temperature and chemical potential, $P = P(T, \mu)$, and is represented by equivalent Eqs.~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}. The second one expresses the pressure as a function of temperature and density, $P = P(T, \rho^*)$, and is represented by equivalent Eqs.~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}. These equations, together with the explicit dependency of density on temperature and chemical potential, $\rho^* = \rho^*(T, \mu)$, Eq.~\eqref{eq:density}, are the basis for the calculation of the thermodynamic response functions in the next Section~\ref{sec:respons_functions}.

@@

Figure~\ref{fig2a} illustrates the dependency of the compressibility $\kappa^*_T$ on the density $\rho^*$ for various values of temperature above the critical one. The dependency of $\kappa^*_T$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig2b}.

@@

Figure~\ref{fig3a} shows the dependency of the pressure coefficient $\beta^*_V$ on the density $\rho^*$ for various values of temperature above the critical one. It is interesting to note that the temperature dependence is very weak in this case, and and multiple isotherms collapse onto the same line and are indistinguishable at the scale of the figure. This is not a surprised behavior, as similar one is observed for the thermal pressure coefficient of the Lennard-Jones fluid~\cite{YigzaweSadus2013} as well. The dependency of $\beta^*_V$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig3b}. The dependency of $\beta^*_V$ on the relative temperature $\tau$ is displayed in Fig.~\ref{fig3c} for a few values of density.

@@

Figure~\ref{fig4a} shows the dependency of the reduced thermal expansion coefficient $\alpha^*_P$ on the density $\rho^*$ for various values of temperature above the critical one. The dependency of $\alpha^*_P$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig4b}.

@@

	\bibitem{YigzaweSadus2013} T.~M. Yigzawe, R.~J. Sadus.
	\href{https://doi.org/10.1063/1.4803855}{Intermolecular interactions and the thermodynamic properties of supercritical fluids.} \textit{J. Chem. Phys.} \textbf{138}, 194502 (2013) [doi: 10.1063/1.4803855]

	\bibitem{Velasco2011} I~Velasco, C.~Rivas, J.~F.~Martinez-Lopez, S.T.~Blanco, S.~Otin, M.~Artal.
	\href{https://pubs.acs.org/doi/abs/10.1021/jp202317n}{Accurate values of some thermodynamic properties for carbon dioxide, ethane, propane, and some binary mixtures.} \textit{J. Phys. Chem. B} \textbf{115}, 8216 (2011) [doi: 10.1021/jp202317n]

	\bibitem{Bulavin2024} L.~A.~Bulavin, Y.~G.~Rudnikov, A.~V.~Chalyi.
	\href{https://pubs.aip.org/aip/adv/article/14/8/085213/3307287}{Contributions to the isothermal compressibility coefficient of water near the temperature of 42$^{\circ}$ C.} \textit{AIP Advances} \textbf{14}, 085213 (2024) [doi: 10.1063/5.0205612]

@@

	\bibitem{BiswasHamann1985} R.~Biswas, D.~R. Hamann.
	\href{https://link.aps.org/doi/10.1103/PhysRevLett.55.2001}{Interatomic
	potentials for silicon structural energies.} \textit{Phys. Rev. Lett.} \textbf{55}, 2001 (1985) [doi: 10.1103/PhysRevLett.55.2001]

	\bibitem{Lim2005} T.-C. Lim. \href{https://acta-arhiv.chem-soc.si/52/52-2-149.htm}{Approximate
	relationships between the generalized Morse and the extended-Rydberg potential energy functions.} \textit{Acta Chim. Slov.} \textbf{52}, 149 (2005)

>>>>

\subsection{Pressure as a function of temperature and density}
In this work, by density we mean the particle number density $\rho = \langle N \rangle / V$.
In the framework of the grand canonical ensemble the average number of particles $\left\langle N \right\rangle$ is found by 
\begin{eqnarray*}
	\langle N \rangle & = & \left(\frac{\partial \ln\Xi}{\partial \beta \mu}\right)_{T,V} 
	\\
	& = & V \left(\frac{\partial P}{\partial \mu}\right)_{T,V}.
\end{eqnarray*}
From this equation it follows
\begin{equation*}
	\frac{\langle N \rangle}{V} = \frac{\langle N \rangle}{v N_v} = \left(\frac{\partial P}{\partial \mu}\right)_{T,V}
\end{equation*}
and thus
\begin{equation}
	\rho^* \equiv \frac{\langle N \rangle}{V} v = v \left(\frac{\partial P}{\partial \mu}\right)_{T,V} 
	= \left(\frac{\partial (Pv\beta)}{\partial (\beta \mu)}\right)_{T,V}.
\end{equation}
The quantity $\rho^*$, on the one hand, is the reduced particle number density, which is the notation commonly used in the literature on simple fluids~\cite{HansenMcDonald2013}. In the context of the cell model, on the other hand, it is the average number of particles per cell.\footnote{In our previous works we denoted the reduced number density by $\bar{n}$. In the current work we switch to more common notation $\rho^*$.}
Taking explicit derivatives, we arrive at
\begin{equation}\label{eq:density}
	\rho^* = \rho^*_c - M + \frac{ \bar \rho_0}{\beta W(0)}.
\end{equation}
The quantity $\rho^*_c$ in the equation \eqref{eq:density} is the critical density~\cite{KozlovskiiDobush2020}
\begin{eqnarray}\label{eq:crit_dens}
	\rho^*_c & = & g_1 - \frac{g_3}{g_4}\left(  g_2 - \frac{1}{2} \frac{g_3^2}{g_4}\right) - \frac{1}{6} \frac{g_3^3}{g_4^2} 
	\nonumber\\
	& = & g_1 - \frac{g_2 g_3}{g_4} + \frac{g_3^3}{3g_4^2}
\end{eqnarray}
Its numerical value for parameters~\eqref{params} is
\begin{equation*}
	\rho^*_c = 0.978.
\end{equation*}
Equations~\eqref{eq:density} and~\eqref{eq:eosMT} jointly define a parametric relationship between pressure and density, with $M$ serving as the parameter. Figure~\ref{fig1} shows the dependence of the reduced pressure $P^*$ on the density $\rho^*$ for various values of the reduced temperature $\tau$, at and above the critical temperature.

This dependence can also be expressed explicitly. To achieve this, we combine the equations~\eqref{eq:density} and~\eqref{eq:ro_M} to express the effective chemical potential $M$ as a function of density and temperature:
\begin{equation}\label{eq:M_nT}
	\bar M = \frac{\rho_{n}}{\beta W(0)} - (\rho^* - \rho^*_c),
\end{equation}
where
\begin{align} \label{eq:ro_nT}
	& \rho_{n} = - 2 \left(\frac{g_3^2 - 2g_2 g_4}{g_4^2} \right)^{\frac{1}{2}} \cos \left( \frac{\alpha_n}{3} + \frac{\pi}{3} \right), \\
	& \alpha_n = \arccos \left[ \left( - \frac{9 g_4^4}{\left( 2 g_2 g_4 - g_3^2\right)^3}\right)^{\frac{1}{2}} (\rho^*_c - \rho^*)\right]. \nonumber 
\end{align}
The notation $\bar{M}$ represents the effective chemical potential as a function of temperature and density $\rho^*$, while $M$ represents the effective chemical potential as a function of temperature and chemical potential~$\mu$.

@@

Thus, in this Section, we presented two forms of the equation of state. The first one expresses the pressure as a function of temperature and chemical potential, $P = P(T, \mu)$, and is represented by equivalent Eqs.~\eqref{eq:eosMT} and~\eqref{eq:eosPTM_reduced}. The second one expresses the pressure as a function of temperature and density, $P = P(T, \rho^*)$, and is represented by equivalent Eqs.~\eqref{eq:eosNT} and~\eqref{eq:eosPTn_reduced}. These equations, together with the explicit dependence of density on temperature and chemical potential, $\rho^* = \rho^*(T, \mu)$, Eq.~\eqref{eq:density}, are the basis for the calculation of the thermodynamic response functions in the next Section~\ref{sec:respons_functions}.

@@

Figure~\ref{fig2a} illustrates the dependence of the compressibility $\kappa^*_T$ on the density $\rho^*$ for various values of temperature above the critical one. The dependence of $\kappa^*_T$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig2b}.

@@

Figure~\ref{fig3a} shows the dependence of the pressure coefficient $\beta^*_V$ on the density $\rho^*$ for various values of temperature above the critical one. It is interesting to note that the temperature dependence is very weak in this case, and multiple isotherms collapse onto the same line and are indistinguishable at the scale of the figure. This is not a surprised behavior, as similar one is observed for the thermal pressure coefficient of the Lennard-Jones fluid~\cite{YigzaweSadus2013} as well. The dependence of $\beta^*_V$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig3b}. The dependence of $\beta^*_V$ on the relative temperature $\tau$ is displayed in Fig.~\ref{fig3c} for a few values of density.

@@

Figure~\ref{fig4a} shows the dependence of the reduced thermal expansion coefficient $\alpha^*_P$ on the density $\rho^*$ for various values of temperature above the critical one. The dependence of $\alpha^*_P$ on the effective chemical potential $M$ is displayed in Fig.~\ref{fig4b}.

@@

	\bibitem{YigzaweSadus2013} T.M. Yigzawe, R.J.~Sadus.
	\href{https://doi.org/10.1063/1.4803855}{Intermolecular interactions and the thermodynamic properties of supercritical fluids.} \textit{J. Chem. Phys.} \textbf{138}, 194502 (2013) [doi: 10.1063/1.4803855]

	\bibitem{Velasco2011} I~Velasco, C.~Rivas, J.F.~Martinez-Lopez, S.T.~Blanco, S.~Otin, M.~Artal.
	\href{https://pubs.acs.org/doi/abs/10.1021/jp202317n}{Accurate values of some thermodynamic properties for carbon dioxide, ethane, propane, and some binary mixtures.} \textit{J. Phys. Chem. B} \textbf{115}, 8216 (2011) [doi: 10.1021/jp202317n]

	\bibitem{Bulavin2024} L.A.~Bulavin, Y.G.~Rudnikov, A.V.~Chalyi.
	\href{https://pubs.aip.org/aip/adv/article/14/8/085213/3307287}{Contributions to the isothermal compressibility coefficient of water near the temperature of 42$^{\circ}$ C.} \textit{AIP Advances} \textbf{14}, 085213 (2024) [doi: 10.1063/5.0205612]

@@

	\bibitem{BiswasHamann1985} R.~Biswas, D.R.~Hamann.
	\href{https://link.aps.org/doi/10.1103/PhysRevLett.55.2001}{Interatomic
	potentials for silicon structural energies.} \textit{Phys. Rev. Lett.} \textbf{55}, 2001 (1985) [doi: 10.1103/PhysRevLett.55.2001]

	\bibitem{Lim2005} T.-C.~Lim. \href{https://acta-arhiv.chem-soc.si/52/52-2-149.htm}{Approximate
	relationships between the generalized Morse and the extended-Rydberg potential energy functions.} \textit{Acta Chim. Slov.} \textbf{52}, 149 (2005)
